---
title: "survival_analysis_insurance_claims.Rmd"
author: "Jiyu Zhang"
date: "2025-04-14"
output: html_document
---
This analysis investigates the survival time to first medical insurance claim for two groups of individuals (Group A and Group B) using:
	•	Kaplan-Meier estimator
	•	Cox proportional hazards model
The dataset is synthetic and includes age, time to first claim, censoring status, and group membership.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load Libraries and Data
```{r libraries}
library(survival)
library(survminer)

# Reproducibility
set.seed(2298)

# Create dataset
Group <- c(rep("A", 5), rep("B", 5))
Age <- c(75, 94, 30, 71, 48, 13, 89, 80, 93, 78)
Time <- c(0.6574794, 2.0665514, 2.0744577, 2.37401, 2.7074722, 
          0.2005577, 0.3497374, 1.6331543, 3.8931681, 4.0719423)
Censor <- c(0, 1, 0, 0, 1, 0, 1, 0, 0, 0)

data <- data.frame(Group, Age, Time, Censor)
data$Group <- as.factor(data$Group)
```

 Kaplan-Meier Estimates
 
```{r km-fit}
km_fit <- survfit(Surv(Time, Censor) ~ Group, data = data)

# Plot
ggsurvplot(km_fit,
           data = data,
           conf.int = TRUE,
           risk.table = TRUE,
           title = "Kaplan-Meier Survival Curve by Group",
           xlab = "Time (Years)",
           ylab = "Survival Probability")
```
	•	Group A tends to reach first claim earlier than Group B.
	•	Censoring is present, and confidence intervals are wider due to small sample size.

Estimate Specific Probabilities
```{r quantities}
# Pr(TA ≤ 2.7) → From KM survival function
Quantity1 <- summary(km_fit, times = 2.7)$surv[1]  # for Group A

# Pr(TB > 4.5) → Not directly defined due to censoring
Quantity2 <- NA  # Cannot estimate beyond observed time
Quantity1
Quantity2
``` 

Cox Proportional Hazards Model
```{r cox-model}
cox_fit <- coxph(Surv(Time, Censor) ~ Group + Age, data = data)
summary(cox_fit)
```
```{r hazard-ratios}
Quantity3 <- coef(cox_fit)[1]  # β1 for Group
Quantity4 <- coef(cox_fit)[2]  # β2 for Age
exp(Quantity3)  # HR for Group
exp(Quantity4)  # HR for Age
```

Likelihood Ratio Test Statistic
```{r lrt}
Quantity6 <- summary(cox_fit)$logtest[1] # Likelihood ratio test statistic
Quantity6
```

Summary of Results
•	Kaplan-Meier Estimate:
	Pr(TA ≤ 2.7) ≈ r round(1 - Quantity1, 2)
	Pr(TB > 4.5) is undefined due to right-censoring beyond 4.07
•	Cox Model Coefficients:
	β1 (Group): r round(Quantity3, 3) → HR ≈ r round(exp(Quantity3), 3)
	β2 (Age): r round(Quantity4, 3) → HR ≈ r round(exp(Quantity4), 3)
•	Likelihood Ratio Test Statistic: r Quantity6
	
Conclusion
•	The analysis suggests Group has a substantial effect on time to first insurance claim, with Group B likely to claim later than Group A.
•	Age has a minor effect, as its hazard ratio is close to 1.
•	Due to small sample size, the statistical power is low; further analysis on a larger dataset is recommended.